The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multi-Phase Forest Inventories

### 2017-10-12 rnw/Calculation of Confidence Intervals.Rnw
update Andreas 12.10.2017(3)

\begin{equation}\label{ci_1phase}
CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2}-1, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})},
                              \hat{Y}+t_{n_{2}-1, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})}\bigg{\rbrack}
\end{equation}

@@

\begin{equation}\label{ci_2phase_global}
CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2}-p, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})},
                              \hat{Y}+t_{n_{2}-p, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})}\bigg{\rbrack}
\end{equation}

@@

\begin{equation}\label{ci_2phase_sae}
CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})},
                              \hat{Y}+t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\var(\hat{Y})}\bigg{\rbrack}
\end{equation}

where $\hat{Y}$ is the point estimate, $\var(\hat{Y})$ is the estimated variance, $1-\alpha$ is the confidence level and $p$ constitutes the number of parameters used in the (full) regression model. In case of cluster-sampling, $n_{2,G}$ is the number of terrestrial clusters (a cluster constitutes the sample unit and comprises multiple sample plots). In \pkg{forestinventory}, the confidence intervals for all estimation methods and estimators can be computed by the S3 generic method \code{confint()}, which requires an estimation object created by either the \code{onephase()}, \code{twophase()} or \code{threephase()} function. For example, the 95\% confidence interval for the small area estimates by the \textit{extpsynth} estimator (Section \ref{sec:twophase_sae}) are calculated by:

>>>>

\begin{equation}\label{ci_1phase}
CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})},
                              \hat{Y}+t_{n_{2}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})}\bigg{\rbrack}
\end{equation}

@@

\begin{equation}\label{ci_2phase_global}
CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2}-p, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})},
                              \hat{Y}+t_{n_{2}-p, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})}\bigg{\rbrack}
\end{equation}

@@

\begin{equation}\label{ci_2phase_sae}
CI_{1-\alpha}(\hat{Y})=\bigg{\lbrack}\hat{Y}-t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})},
                              \hat{Y}+t_{n_{2,G}-1, 1-\frac{\alpha}{2}}\sqrt{\hat{\var}(\hat{Y})}\bigg{\rbrack}
\end{equation}

where $\hat{Y}$ is the point estimate, $\hat{\var}(\hat{Y})$ is the estimated variance, $1-\alpha$ is the confidence level and $p$ constitutes the number of parameters used in the (full) regression model. In case of cluster-sampling, $n_{2,G}$ is the number of terrestrial clusters (a cluster constitutes the sample unit and comprises multiple sample plots). In \pkg{forestinventory}, the confidence intervals for all estimation methods and estimators can be computed by the S3 generic method \code{confint()}, which requires an estimation object created by either the \code{onephase()}, \code{twophase()} or \code{threephase()} function. For example, the 95\% confidence interval for the small area estimates by the \textit{extpsynth} estimator (Section \ref{sec:twophase_sae}) are calculated by:

### 2017-10-12 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 12.10.2017(3)

We will now draw our attention to the calculation of the explanatory variables from the auxiliary data for both the non-exhaustive and exhaustive cases. Figure \ref{fig:exh_nexh_and_boundweights_b} depicts how the \textbf{non-exhaustive} case often looks like in practice: a regular terrestrial grid $s_2$ is given by a terrestrial inventory (the points surrounded by dotted circles in Fig. \ref{fig:exh_nexh_and_boundweights_b}) and densified to a larger sample $s_1$ (the points). For every point $x$, each explanatory variable in the vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^t$ is calculated using a defined spatial extent of auxiliary information around that point called the \textit{support} (the dark green square tiles). We emphasize that the value of the explanatory variables for $\pmb{Z}(x)$ are associated with the point whereas the support is the spatial extent of the auxiliary information used to calculate those values. So far this is in perfect agreement with the presented theory of the non-exhaustive estimator, except for using regular grids rather than randomly placed sample points. The \pkg{forestinventory} package calculates the empirical mean of $\pmb{Z}(x)$ automatically from the input data frame using $\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.\par

@@

One possibility to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Fig. \ref{fig:exh_nexh_and_boundweights_a}, where the spatial arrangement of the supports (the dark green tiles) are tessellated to form a perfect partition over the inventory area in order for all of the wall-to-wall auxiliary information to be exploited. Is has to be noted that this setup would allow for a perfect calculation of the exact mean $\bar{\pmb{Z}}$ in the \textit{finite} population approach, i.e. deriving $\pmb{Z}(x)$ for the finite population of supports that are considered the sampling units. While in the infinite population approach this implementation probably does not produce the true exact mean $\bar{\pmb{Z}}$, $n_1$ is still expected to be reasonably large for the difference to be considered negligible as long as the size of the supports are not unreasonably large. However, the perfect tessellation implementation can also impose drawbacks. One is that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ and the support size. Since in practice the support size should ideally be chosen to achieve a best possible explanatory power of the regression model (thus minimizing the residual variation) a perfect tesselation might often not be feasible. In the infinite population frame however, the supports are allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ to get a negligibly close approximation of $\bar{\pmb{Z}}$. With this respect, the infinite population approach provides more flexibility than the finite approach.

>>>>

We will now draw our attention to the calculation of the explanatory variables from the auxiliary data for both the non-exhaustive and exhaustive cases. Figure \ref{fig:exh_nexh_and_boundweights_b} depicts how the \textbf{non-exhaustive} case often looks like in practice: a regular terrestrial grid $s_2$ is given by a terrestrial inventory (the points surrounded by dotted circles in Fig. \ref{fig:exh_nexh_and_boundweights_b}) and densified to a larger sample $s_1$ (the points). For every point $x$, each explanatory variable in the vector $\pmb{Z}(x)=(z(x)_1, z(x)_2,...,z(x)_p)^t$ is calculated using a defined spatial extent of auxiliary information around that point called the \textit{support} (the dark green square tiles). We emphasize that the value of the explanatory variables for $\pmb{Z}(x)$ are associated with the sample point whereas the support is the spatial extent of the auxiliary information used to calculate those values. So far this is in perfect agreement with the presented theory of the non-exhaustive estimator, except for using regular grids rather than randomly placed sample points. The \pkg{forestinventory} package calculates the empirical mean of $\pmb{Z}(x)$ automatically from the input data frame using $\hat{\bar{\pmb{Z}}}=\frac{1}{n_{1}}\sum_{x\in{s_{1}}}\pmb{Z}(x)$.\par

@@

One implementation to approximate the exact mean $\bar{\pmb{Z}}$ is shown in Fig. \ref{fig:exh_nexh_and_boundweights_a}, where the spatial arrangement of the supports (the dark green tiles) are tessellated to form a perfect partition over the inventory area in order for all of the wall-to-wall auxiliary information to be exploited. Is has to be noted that this setup would allow for a perfect calculation of the exact mean $\bar{\pmb{Z}}$ in the \textit{finite} population approach, i.e. deriving $\pmb{Z}(x)$ for the finite population of supports that are considered the sampling units. While in the infinite population approach this implementation probably does not produce the true exact mean $\bar{\pmb{Z}}$, $n_1$ is still expected to be reasonably large for the difference to be considered negligible as long as the size of the supports are not unreasonably large. However, the perfect tessellation implementation can also impose drawbacks. One is that a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ and the support size. Since in practice the support size should ideally be chosen to achieve a best possible explanatory power of the regression model (thus minimizing the residual variation) a perfect tesselation might often not be feasible. In the infinite population frame, the supports are allowed to overlap if this seems necessary to acquire a sufficiently large sample $n_1$ to get a negligibly close approximation of $\bar{\pmb{Z}}$. With this respect, the infinite population approach provides more flexibility than the finite approach.
